The Galactic Midplane is Not a Plane: Implications for {\em Gaia}

### 2019-04-10 paper/ms.tex
Rerun to check numbers, minor changes to ttext

Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
midplane our imaginary observer would determine as a function of azimuth for
each galaxy, using their local ``solar neighborhood'' (the cylinder defined
above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
shown as the dashed-line error bars. The middle $90\%$ of midplane values across the Solar circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
three cases the midplane therefore varies by more than $\pm 100\,\pc$
depending on the azimuth along the Solar circle; in the third (\mm{}, which
has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
variation is more like $\pm 50\,\pc$.

@@

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine error bars
(95\uth and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

>>>>

Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
midplane our imaginary observer would determine as a function of azimuth for
each galaxy, using their local ``solar neighborhood'' (the cylinder defined
above). The $1\,\sigma$ error from the bootstrap procedure is
shown as the dashed-line error bars. The middle $90\%$ of midplane values across the Solar circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
three cases the midplane therefore varies by more than $\pm 100\,\pc$
depending on the azimuth along the Solar circle; in the third (\mm{}, which
has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
variation is more like $\pm 50\,\pc$.

@@

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $error bars, which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}
